---
title: "Lesson 2 - Chapter 1 - 1, 2, 3"
author: "Samantha Chinn"
date: "1/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, results = "show", fig.show = "show")
library(tidyverse)
library(ma391chinn)
```

### Problem 1 - Car
#### (a) What amount of rebate will maximize profit?
##### Step 1: 

* p = profit on sale model ($1500/car)
* r = rebate $/car
* x = rebate scale factor ($100/car)
* y = sale increase based on rebate (0.15 = 15%)
* n = number of cars sold
* P = total profit

##### Step 2: Select the modeling approach
Single Variable Optimization

##### Step 3: Formulate the model
f = function(r){(1500-r)*(1+0.15(r/100))}
df = function(r){fprime(f,r)}

##### Step 4: Solve the model
r = seq(0,20)
ans = bisection(df,0,20)
print(ans)
print(f(ans))






### Problem 2 - Pig

* x = time (days)

```{r}
cost = function(x){(0.65-0.01*x)*(200+7*x)-(0.60*x)}
dCost = function(x){fprime(cost,x)}
x = seq(0,20)
ans = bisection(dCost,0,20)
print(ans)
print(cost(ans))
```
At x = 13.9 the cost is $143.58.




### Problem 3 - Pig
#### (a)
```{r}
pnew = function(x){0.65-0.01*x+0.00004*x^2}
plot(cost,col="red")
plot(pnew,col="blue")
```




